---
title: "Differentialy expression in Normal Tissue by LN group"
author: "Sara Gosline"
date: "7/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('loadHNSCCdata.R')
```


## Dive into lymph node status to investigate possible data separation

To build a model we want to break down patient samples into various classes based on their lymph node status. 

Let's first start by seeing how many patients have positive lymph nodes:

```{r plot of LN counts, warn=FALSE}
       ggplot(clin.dat,aes(x=ln_examined,y=positive_by_he,col=staging))+geom_point()
```
       
It seems that there is a large number that have 0 positive lymph nodes, a group that has just 1, and a group that has 2 or more.
```{r plots of groups}

       
       
ggplot(red.dat,aes(x=lnCounts,fill=staging))+geom_bar(position='dodge')
    
```

## Group samples and get differential expression

Now we can compute proteins that are differentially expressed between LN classes.

```{r differential expression}
library(limma)

#create data matrix of samples and collect patients
protMat<-normProts%>%
        pivot_wider(values_from='inferredNormAbund',names_from='patient')%>%
        tibble::column_to_rownames('Gene')

zpats=red.dat%>%subset(lnCounts=='zeroLN')%>%
        select(patient)%>%
        subset(patient%in%colnames(protMat))

opats=red.dat%>%subset(lnCounts=='oneOrTwo')%>%
        select(patient)%>%
        subset(patient%in%colnames(protMat))

mpats=red.dat%>%subset(lnCounts=='moreThanTwo')%>%
        select(patient)%>%
        subset(patient%in%colnames(protMat))

doLim<-function(dat,sampleIDs.group1,sampleIDs.group2){
  fac <- factor(rep(c(2,1), c(length(sampleIDs.group2), length(sampleIDs.group1))))
  design <- model.matrix(~fac)
  fit <- lmFit(dat[,c(sampleIDs.group2, sampleIDs.group1)], design)
  fit <- eBayes(fit)
  print(topTable(fit, coef=2))
  res <- topTable(fit, coef=2, number=Inf, sort.by="none")
  res <- data.frame(featureID=rownames(res), res, stringsAsFactors = F)
  return(res)
}

oneVsZero=doLim(protMat,opats$patient,zpats$patient)
moreVsOne=doLim(protMat,mpats$patient,opats$patient)
moreVsZero=doLim(protMat,mpats$patient,zpats$patient)

allVsZero=doLim(protMat,c(opats$patient,mpats$patient),zpats$patient)
##now 
```

So we have some proteins!! Only 1 is differentially expressed. Let's try some GSEA analysis instead to see if we can find biological processes enriched.

## Biological pathway enrichment

Let's compute the pathways (NCI Pathways) enriched upon proteins that are different between zero and 1/2 LN affected

```{r biological pathways,warning=FALSE}
library(leapr)

data("ncipid")
data('krbpaths')
z.en <- leapR(geneset=ncipid,enrichment_method='enrichment_in_order',datamatrix=oneVsZero,primary_columns='logFC')%>%
        subset(BH_pvalue<0.05)

print(z.en)

```

Now let's look at pathways that go from 1/2 LN to more than 1/2

```{r biological pathways more,warning=FALSE}

m.en <-leapR(geneset=ncipid,enrichment_method='enrichment_in_order',datamatrix=moreVsOne,primary_columns='logFC')%>%
        subset(BH_pvalue<0.25)

print(m.en)

```

## Plot changing proteins

Now let's see if we can plot the patient samples by the proteins involved


```{r, plot samples,warning=FALSE}
zprots=subset(oneVsZero,P.Value<0.01)%>%select(featureID)
mprots=subset(moreVsOne,P.Value<0.01)%>%select(featureID)

redProts <- fullProts%>%
  subset(Gene%in%c(zprots$featureID,mprots$featureID))

##let's first do a PCA plotted with positive lns
tumAbund <- pcaFromDf(redProts,'inferredNormAbund')%>%
  left_join(clin.dat,by='patient')

##all proteins
tumAbund%>%left_join(select(red.dat,c(patient,lnCounts)),by='patient')%>%
ggplot()+
  geom_point(aes(x=PC1,y=PC2,col=lnCounts,size=positive_by_he))+
  ggtitle('Selected proteins by normal expression')

```
