---
title: "Differentialy altered imaging features by LN group"
author: "Sara Gosline"
date: "7/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source('loadHNSCCdata.R')
```

## Evaluate clustering of samples by imaging features

How do the tumor samples cluster by imaging features alone?

```{r cluster by image, warning=FALSE}
##let's first do a PCA plotted with positive lns
imgVals <- pcaFromDf(rename(imgFeat,Gene='Feature_Type'),'value')%>%
  left_join(clin.dat,by='patient')%>%
  left_join(select(red.dat,c(patient,lnCounts)),by='patient')

##all proteins
ggplot(imgVals)+
  geom_point(aes(x=PC1,y=PC2,size=ln_examined,col=lnCounts))+
  ggtitle('All tumors by image features')+scale_color_manual(values=pal)


##all proteins
ggplot(imgVals)+
  geom_point(aes(x=PC2,y=PC3,size=ln_examined,col=lnCounts))+
  ggtitle('All tumors by image features (PC2 and 3)')+scale_color_manual(values=pal)


```

Those two outliers are not great - might throw things off a bit, so let's remove them.

```{r remove outliers,warning=FALSE}


outliers <- imgVals%>%subset(PC1<(-10e9)|PC2<(-2e9))%>%
  select(patient)


imgFeat <- imgFeat%>%
  subset(patient!=outliers$patient)

imgVals <- pcaFromDf(rename(imgFeat,Gene='Feature_Type'),'value')%>%
  left_join(clin.dat,by='patient')%>%
  left_join(select(red.dat,c(patient,lnCounts)),by='patient')

##all proteins
ggplot(imgVals)+
  geom_point(aes(x=PC1,y=PC2,size=ln_examined,col=lnCounts))+
  ggtitle('All tumors by image features - outliers removed')+scale_color_manual(values=pal)


##all proteins
ggplot(imgVals)+
  geom_point(aes(x=PC2,y=PC3,size=ln_examined,col=lnCounts))+
  ggtitle('All tumors by image features (PC2 and 3) outliers removed')+scale_color_manual(values=pal)


```


## Visualize imaging features in heatmap

Let's see what the imaging features can cluster without any real guidance

```{r heatmap of imaging, warning=FALSE}
library(pheatmap)
imgMat<-imgFeat%>%
  select(-Feature_Class)%>%
        pivot_wider(values_from='value',names_from='patient',values_fn=(value=mean))%>%
        tibble::column_to_rownames('Feature_Type')

mostVar <- apply(log10(imgMat),1,var)%>%
  sort(decreasing=TRUE)


ann <- red.dat%>%
  tibble::remove_rownames()%>%
  tibble::column_to_rownames('patient')

pheatmap(log10(imgMat[names(mostVar)[1:25],]), annotation_col = ann, cellwidth = 10,cellheight=10,filename='variableImgFeatures.pdf')

```

## Group samples and get differential expression

Now we can compute proteins that are differentially expressed between LN classes.

```{r differential expression}
library(limma)

#create data matrix of samples and collect patients
imgMat<-imgFeat%>%
  select(-Feature_Class)%>%
        pivot_wider(values_from='value',names_from='patient',values_fn=(value=mean))%>%
        tibble::column_to_rownames('Feature_Type')

zpats=red.dat%>%subset(lnCounts=='zeroLN')%>%
        select(patient)%>%
        subset(patient%in%colnames(imgMat))

opats=red.dat%>%subset(lnCounts=='oneOrTwo')%>%
        select(patient)%>%
        subset(patient%in%colnames(imgMat))

mpats=red.dat%>%subset(lnCounts=='moreThanTwo')%>%
        select(patient)%>%
        subset(patient%in%colnames(imgMat))

doLim<-function(dat,sampleIDs.group1,sampleIDs.group2){
  fac <- factor(rep(c(2,1), c(length(sampleIDs.group2), length(sampleIDs.group1))))
  design <- model.matrix(~fac)
  fit <- lmFit(dat[,c(sampleIDs.group2, sampleIDs.group1)], design)
  fit <- eBayes(fit)
  res <- topTable(fit, coef=2, number=Inf, sort.by="none")
  res <- data.frame(featureID=rownames(res), res, stringsAsFactors = F)
  return(res)
}

oneVsZero=doLim(imgMat,opats$patient,zpats$patient)
moreVsOne=doLim(imgMat,mpats$patient,opats$patient)
moreVsZero=doLim(imgMat,mpats$patient,zpats$patient)

allVsZero=doLim(imgMat,c(opats$patient,mpats$patient),zpats$patient)
##now
pthresh=0.05
combinedDF<-rbind(data.frame(condition='oneVsZero',subset(oneVsZero,P.Value<pthresh)),
                  data.frame(condition='moreVsOne',subset(moreVsOne,P.Value<pthresh)),
                  data.frame(condition='moreVsZero',subset(moreVsZero,P.Value<pthresh)))
```

So we have some proteins!! Only 1 is differentially expressed. Let's try some GSEA analysis instead to see if we can find biological processes enriched.




## Plot changing proteins

Now let's see if we can plot the patient samples by the proteins involved


```{r, plot samples,warning=FALSE}
zprots=subset(oneVsZero,P.Value<0.05)%>%select(featureID)
mprots=subset(moreVsOne,P.Value<0.05)%>%select(featureID)

redProts <- imgFeat%>%
  subset(Feature_Type%in%c(zprots$featureID,mprots$featureID))

##let's first do a PCA plotted with positive lns
tumAbund <- pcaFromDf(rename(redProts,Gene='Feature_Type'),'value')%>%
  left_join(clin.dat,by='patient')

##all proteins
tumAbund%>%left_join(select(red.dat,c(patient,lnCounts)),by='patient')%>%
ggplot()+
  geom_point(aes(x=PC1,y=PC2,col=lnCounts,size=positive_by_he))+
  ggtitle('Patients by selected imaging features')+scale_color_manual(values=pal)

```

Since there are so few features we can plot each one. 

```{r plot img features}

ggplot(left_join(redProts,red.dat),aes(fill=lnCounts,y=log10(value),x=Feature_Type))+geom_boxplot()+guides(x=guide_axis(angle=45))+scale_fill_manual(values=pal)

```