---
title: "Build up the Bayesian network"
author: "Song Feng"
date: "08/07/2021"
output: html_notebook
---

loading the package

```{r}
library(ggplot2)
library(gplots)
library(bnlearn)
```

Load the data

```{r}
source('loadHNSCCdata.R')
library(reshape)
tumProtTable <- cast(tumProts, Gene~patient, mean)
rownames(tumProtTable) <- paste0(tumProtTable$Gene, "_T")
tumProtTable$Gene <- NULL
normProtTable <- cast(normProts, Gene~patient, mean)
rownames(normProtTable) <- paste0(normProtTable$Gene, "_N")
normProtTable$Gene <- NULL
imgFeatTable <- cast(imgFeat, Feature_Type~patient, mean)
rownames(imgFeatTable) <- imgFeatTable$Feature_Type
imgFeatTable$Feature_Type <- NULL

lnTableAll <- red.dat
lnTableAll$affected <- abs(unclass(factor(lnTableAll$lnSpread)) - 2)
lnTableAll$NoLN <- abs(unclass(factor(lnTableAll$lnCounts)) - 3)
lnTable <- as.data.frame(lnTableAll[ ,c("patient", "affected", "NoLN")])
lnAfTable <- as.data.frame(lnTableAll[ ,c("patient", "affected")])
lnNoTable <- as.data.frame(lnTableAll[ ,c("patient", "NoLN")])

rownames(lnTable) <- lnTable$patient
rownames(lnAfTable) <- lnTable$patient 
rownames(lnNoTable) <- lnTable$patient
lnTable$patient <- NULL 
lnAfTable$patient <- NULL 
lnNoTable$patient <- NULL
lnTableF <- as.data.frame(t(lnTable))
lnAfTableF <- as.data.frame(t(lnAfTable))
lnNoTableF <- as.data.frame(t(lnNoTable))
```


```{r}
load("./analysis_SF/allCorrs.RData")
```

```{r}
features <- c("Complexity", "InterquartileRange", "LeastAxisLength", "MajorAxisLength", "Maximum2DDiameterRow", "Maximum2DDiameterSlice", "MeshVolume", "SizeZoneNonUniformity", "SmallDependenceHighGrayLevelEmphasis", "SurfaceArea", "TotalEnergy", "VoxelVolume")

imgFeatTableF <- imgFeatTable[features, ]
imgFeatTableF
```


```{r}
load("./analysis_SF/allRawCorrs.RData")

filterCorr <- function(corTable, th = 0.6) {
  return(as.data.frame(corTable[,colSums((corTable > th) | (corTable < -th))> 0]))
}

corrIfTpF <- filterCorr(corrIfTp[features,], 0.5) # 0.5
tumSelected <- paste0(colnames(corrIfTpF), "_T")
print(length(tumSelected))
tumProtTableF <- tumProtTable[tumSelected,]
corrIfNpF <- filterCorr(corrIfNp[features,], 0.6) # 0.6
normSelected <- paste0(colnames(corrIfNpF), "_N")
print(length(normSelected))
normProtTableF <- normProtTable[normSelected,]
```

```{r}
tumProtTableF
```

```{r}
normProtTableF
```



```{r}
col_names <- Reduce(intersect, list(colnames(imgFeatTableF), colnames(tumProtTableF), colnames(normProtTableF), colnames(lnTableF)))
imgFeatProtTable <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], tumProtTableF[, col_names], normProtTableF[, col_names], lnTableF[, col_names]))))
imgFeatProtTable
```


```{r}
col_names <- Reduce(intersect, list(colnames(imgFeatTableF), colnames(tumProtTableF), colnames(normProtTableF)))
imgFeatProtTable <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], tumProtTableF[, col_names], normProtTableF[, col_names]))))
imgFeatProtTable
```


```{r}
rho = cor(imgFeatProtTable)
heatmap.2(rho, scale = "none", trace = "none", revC = TRUE)#, breaks = palette.breaks)
```

```{r}
ug = empty.graph(colnames(rho))
amat(ug) = (rho > 0.6) + 0L - diag(1L, nrow(rho))
graphviz.plot(ug, layout = "fdp", shape = "ellipse")
```

```{r}
bl <- rbind(
    set2blacklist(c(rownames(tumProtTableF), rownames(normProtTableF))), 
    set2blacklist(rownames(lnTableF)), 
    set2blacklist(rownames(imgFeatTableF)), 
    tiers2blacklist(list(c(rownames(tumProtTableF), rownames(normProtTableF)),  
                         rownames(imgFeatTableF))), 
    tiers2blacklist(list(rownames(lnTableF), rownames(imgFeatTableF))),
    tiers2blacklist(list(c(rownames(tumProtTableF), rownames(normProtTableF)), 
                         rownames(lnTableF)))#,
    # tiers2blacklist(list(c(rownames(tumProtTableF), rownames(normProtTableF)), 
    #                      rownames(imgFeatTableF)))
    )
wl <- rbind(
    tiers2blacklist(list(rownames(lnTableF), 
                         c(rownames(tumProtTableF), rownames(normProtTableF)))),
    # tiers2blacklist(list(rownames(imgFeatTableF), 
    #                      c(rownames(tumProtTableF), rownames(normProtTableF)))),
    tiers2blacklist(list(rownames(imgFeatTableF), rownames(lnTableF)))
    )
```


```{r}
bl <- rbind(
    # set2blacklist(c(rownames(tumProtTableF), rownames(normProtTableF))),
    # set2blacklist(rownames(lnTableF)),
    # set2blacklist(rownames(imgFeatTableF)), 
    # tiers2blacklist(list(c(rownames(tumProtTableF), rownames(normProtTableF)),  
    #                      rownames(imgFeatTableF))), 
    # tiers2blacklist(list(rownames(imgFeatTableF), rownames(lnTableF))),
    # tiers2blacklist(list(c(rownames(tumProtTableF), rownames(normProtTableF)),
    #                      rownames(lnTableF)))#,
    # tiers2blacklist(list(c(rownames(tumProtTableF), rownames(normProtTableF)), 
    #                      rownames(imgFeatTableF)))
    )
wl <- rbind(
    # tiers2blacklist(list(rownames(lnTableF), 
    #                      c(rownames(tumProtTableF), rownames(normProtTableF)))),
    # tiers2blacklist(list(rownames(imgFeatTableF), 
    #                      c(rownames(tumProtTableF), rownames(normProtTableF)))),
    # tiers2blacklist(list(rownames(imgFeatTableF), rownames(lnTableF)))
    )
```


```{r}
str.raw = boot.strength(imgFeatProtTable, R = 500, algorithm = "hc", algorithm.args = list(blacklist = bl, whitelist = wl))
attr(str.raw, "threshold")
avg.raw.full = averaged.network(str.raw)
```

```{r}
strength.plot(avg.raw.full, str.raw, shape = "rectangle", layout = "dot")#, highlight = list(arcs = wl))
```

```{r}
strength.plot(avg.raw.full, str.raw, threshold = 0.6, shape = "rectangle", layout = "dot")#, highlight = list(arcs = wl))
```


```{r}

avg.raw.full$learning$whitelist = wl
avg.raw.full$learning$blacklist = bl
nrow(undirected.arcs(cpdag(avg.raw.full, wlbl = TRUE)))
```

```{r}
nrow(str.raw[with(str.raw, strength > 0.50 & direction > 0.50), ])
```

```{r}
nrow(str.raw[with(str.raw, strength > 0.85 & direction > 0.50), ])
```


```{r}
min(str.raw[with(str.raw, strength > 0.50 & direction > 0.50), "direction"])
```


```{r}
avg.raw.simpler = averaged.network(str.raw, threshold = 0.8)
g <- Rgraphviz::layoutGraph(bnlearn::as.graphNEL(avg.raw.simpler))
graph::nodeRenderInfo(g) <- list(fontsize=100)
Rgraphviz::renderGraph(g)
strength.plot(avg.raw.simpler, str.raw, shape = "ellipse")#, highlight = list(arcs = wl))
```

```{r}
dag = hc(imgFeatProtTable, whitelist = wl, blacklist = bl)
dag
```

```{r}
str.raw = boot.strength(imgFeatProtTable, R = 100, algorithm = "hc", algorithm.args = list(blacklist = bl, whitelist = wl))
attr(str.raw, "threshold")
avg.raw.full = averaged.network(str.raw)
```

```{r}
strength.plot(avg.raw.full, str.raw, shape = "ellipse")#, highlight = list(arcs = wl))
```

```{r}
xval = bn.cv(imgFeatProtTable, bn = "hc", algorithm.args = list(blacklist = bl, whitelist = wl), loss = "cor-lw", loss.args = list(target = "affected", n = 200), runs = 10)
```

# Compare different models and features

```{r}
col_names <- Reduce(intersect, list(colnames(imgFeatTableF), colnames(tumProtTableF), colnames(normProtTableF), colnames(lnTableF)))

img_prot_af <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], tumProtTableF[, col_names], normProtTableF[, col_names], lnAfTableF[, col_names]))))
img_tum_af <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], tumProtTableF[, col_names], lnAfTableF[, col_names]))))
img_norm_af <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], normProtTableF[, col_names], lnAfTableF[, col_names]))))
img_af <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], lnAfTableF[, col_names]))))
prot_af <- data.frame(t(data.frame(rbind(tumProtTableF[, col_names], normProtTableF[, col_names], lnAfTableF[, col_names]))))

img_prot_no <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], tumProtTableF[, col_names], normProtTableF[, col_names], lnNoTableF[, col_names]))))
img_no <- data.frame(t(data.frame(rbind(imgFeatTableF[, col_names], lnNoTableF[, col_names]))))
prot_no <- data.frame(t(data.frame(rbind(tumProtTableF[, col_names], normProtTableF[, col_names], lnNoTableF[, col_names]))))
```

```{r}
bn_img_af <- empty.graph(c(rownames(imgFeatTableF), rownames(lnAfTableF)))
wl_img_af <- tiers2blacklist(list(rownames(lnAfTableF), rownames(imgFeatTableF)))
arcs(bn_img_af) <- wl_img_af

bn_prot_af <- empty.graph(c(rownames(tumProtTableF), rownames(normProtTableF), rownames(lnAfTableF)))
wl_prot_af <- tiers2blacklist(list(rownames(lnAfTableF), c(rownames(tumProtTableF), rownames(normProtTableF))))
arcs(bn_prot_af) <- wl_prot_af

bn_prot_img_af <- empty.graph(c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(normProtTableF), rownames(lnAfTableF)))
wl_prot_img_af <- rbind(
  tiers2blacklist(list(rownames(lnAfTableF), rownames(imgFeatTableF))),
  tiers2blacklist(list(rownames(imgFeatTableF), c(rownames(tumProtTableF), rownames(normProtTableF))))
)
arcs(bn_prot_img_af) <- wl_prot_img_af

bn_tum_img_af <- empty.graph(c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(lnAfTableF)))
wl_tum_img_af <- rbind(
  tiers2blacklist(list(rownames(lnAfTableF), rownames(imgFeatTableF))),
  tiers2blacklist(list(rownames(imgFeatTableF), c(rownames(tumProtTableF))))
)
arcs(bn_tum_img_af) <- wl_tum_img_af

bn_norm_img_af <- empty.graph(c(rownames(imgFeatTableF), rownames(normProtTableF), rownames(lnAfTableF)))
wl_norm_img_af <- rbind(
  tiers2blacklist(list(rownames(lnAfTableF), rownames(imgFeatTableF))),
  tiers2blacklist(list(rownames(imgFeatTableF), rownames(normProtTableF)))
)
arcs(bn_norm_img_af) <- wl_norm_img_af

bn_prot_img_id_af <- empty.graph(c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(normProtTableF), rownames(lnAfTableF)))
wl_prot_img_id_af <- tiers2blacklist(list(rownames(lnAfTableF), c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(normProtTableF))))
arcs(bn_prot_img_id_af) <- wl_prot_img_id_af
```

```{r}
bn_img_no <- empty.graph(c(rownames(imgFeatTableF), rownames(lnNoTableF)))
wl_img_no <- tiers2blacklist(list(rownames(lnNoTableF), rownames(imgFeatTableF)))
arcs(bn_img_no) <- wl_img_no

bn_prot_no <- empty.graph(c(rownames(tumProtTableF), rownames(normProtTableF), rownames(lnNoTableF)))
wl_prot_no <- tiers2blacklist(list(rownames(lnNoTableF), c(rownames(tumProtTableF), rownames(normProtTableF))))
arcs(bn_prot_no) <- wl_prot_no

bn_prot_img_no <- empty.graph(c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(normProtTableF), rownames(lnNoTableF)))
wl_prot_img_no <- rbind(
  tiers2blacklist(list(rownames(lnNoTableF), rownames(imgFeatTableF))),
  tiers2blacklist(list(rownames(imgFeatTableF), c(rownames(tumProtTableF), rownames(normProtTableF))))
)
arcs(bn_prot_img_no) <- wl_prot_img_no

bn_prot_img_id_no <- empty.graph(c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(normProtTableF), rownames(lnNoTableF)))
wl_prot_img_id_no <- tiers2blacklist(list(rownames(lnNoTableF), c(rownames(imgFeatTableF), rownames(tumProtTableF), rownames(normProtTableF))))
arcs(bn_prot_img_id_no) <- wl_prot_img_id_no
```


```{r}
graphviz.plot(bn_img_af)
graphviz.plot(bn_prot_af)
graphviz.plot(bn_prot_img_af)
graphviz.plot(bn_prot_img_id_af)
graphviz.plot(bn_tum_img_af)
graphviz.plot(bn_norm_img_af)
```



```{r}
img_af_cv <- bn.cv(img_af, bn = bn_img_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
prot_af_cv <- bn.cv(prot_af, bn = bn_prot_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_prot_af_cv <- bn.cv(img_prot_af, bn = bn_prot_img_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_tum_af_cv <- bn.cv(img_tum_af, bn = bn_tum_img_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_norm_af_cv <- bn.cv(img_norm_af, bn = bn_norm_img_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_prot_id_af_cv <- bn.cv(img_prot_af, bn = bn_prot_img_id_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
```

```{r}
plot(img_af_cv, prot_af_cv, img_prot_af_cv, img_tum_af_cv, img_norm_af_cv, img_prot_id_af_cv, xlab = c("Image only", "Proteomics only", "Proteomics -> Image", "Tum prot -> Image", "Norm prot -> Image", "Image + Proteomics"))
plot(img_af_cv, img_prot_af_cv, xlab = c("Image only", "Proteomics -> Image"))
plot(img_af_cv, img_prot_af_cv, img_tum_af_cv, img_norm_af_cv, xlab = c("Image only", "Proteomics -> Image", "Tum prot -> Image", "Norm prot -> Image"))
```


```{r}
save(img_af_cv, prot_af_cv, img_prot_af_cv, img_tum_af_cv, img_norm_af_cv, img_prot_id_af_cv, file = "./analysis_SF/comparisons.RData")
```


valid learning algorithm(s) are "gs" (Grow-Shrink), "iamb" (IAMB), "fast.iamb" (Fast-IAMB), "inter.iamb" (Inter-IAMB), "iamb.fdr" (IAMB-FDR), "pc.stable" (PC (Stable)), "mmpc" (Max-Min Parent Children), "si.hiton.pc" (Semi-Interleaved HITON-PC), "hpc" (Hybrid Parents and Children), "hc" (Hill-Climbing), "tabu" (Tabu Search), "rsmax2" (Two-Phase Restricted Maximization), "mmhc" (Max-Min Hill-Climbing), "h2pc" (Hybrid^2 Parent Children), "chow.liu" (Chow-Liu), "aracne" (ARACNE), "naive.bayes" (Naive Bayes Classifier), "tree.bayes" (TAN Bayes Classifier). See ?bnlearn-package for details.


```{r}
img_af_ho <- bn.cv(img_af, bn = bn_img_af, runs = 10, method = "hold-out", k = 10, m = 10, loss = "mse-lw", loss.args = list(target = "affected"))
prot_af_ho <- bn.cv(prot_af, bn = bn_prot_af, runs = 10, method = "hold-out", k = 10, m = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_prot_af_ho <- bn.cv(img_prot_af, bn = bn_prot_img_af, runs = 10, method = "hold-out", k = 10, m = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_tum_af_ho <- bn.cv(img_tum_af, bn = bn_tum_img_af, runs = 10, method = "hold-out", k = 10, m = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_norm_af_ho <- bn.cv(img_norm_af, bn = bn_norm_img_af, runs = 10, method = "hold-out", k = 10, m = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_prot_id_af_ho <- bn.cv(img_prot_af, bn = bn_prot_img_id_af, runs = 10, method = "hold-out", k = 10, m = 10, loss = "mse-lw", loss.args = list(target = "affected"))
plot(img_af_ho, prot_af_ho, img_prot_af_ho, img_tum_af_ho, img_norm_af_ho, img_prot_id_af_ho, xlab = c("Image only", "Proteomics only", "Proteomics -> Image", "Tum prot -> Image", "Norm prot -> Image", "Image + Proteomics"))
plot(img_af_ho, img_prot_af_ho, xlab = c("Image only", "Proteomics -> Image"))
plot(img_af_ho, img_prot_af_ho, img_tum_af_ho, img_norm_af_ho, xlab = c("Image only", "Proteomics -> Image", "Tum prot -> Image", "Norm prot -> Image"))
```



```{r}
img_af_cv <- bn.cv(img_af, bn = bn_img_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
prot_af_cv <- bn.cv(prot_af, bn = bn_prot_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_prot_af_cv <- bn.cv(img_prot_af, bn = bn_prot_img_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
img_prot_id_af_cv <- bn.cv(img_prot_af, bn = bn_prot_img_id_af, runs = 10, loss = "mse-lw", loss.args = list(target = "affected"))
plot(img_af_cv, prot_af_cv, img_prot_af_cv, img_prot_id_af_cv, xlab = c("Image only", "Proteomics only", "Proteomics -> Image", "Image + Proteomics"))
plot(img_af_cv, img_prot_af_cv, xlab = c("Image only", "Proteomics -> Image"))
```


```{r}
img_af_cv <- bn.cv(img_af, bn = bn_img_af, runs = 10, loss = "mse", loss.args = list(target = "affected"))
prot_af_cv <- bn.cv(prot_af, bn = bn_prot_af, runs = 10, loss = "mse", loss.args = list(target = "affected"))
img_prot_af_cv <- bn.cv(img_prot_af, bn = bn_prot_img_af, runs = 10, loss = "mse", loss.args = list(target = "affected"))
img_prot_id_af_cv <- bn.cv(img_prot_af, bn = bn_prot_img_id_af, runs = 10, loss = "mse", loss.args = list(target = "affected"))
plot(img_af_cv, prot_af_cv, img_prot_af_cv, img_prot_id_af_cv, xlab = c("Image only", "Proteomics only", "Proteomics -> Image", "Image + Proteomics"))
plot(img_af_cv, img_prot_af_cv, xlab = c("Image only", "Proteomics -> Image"))
```


```{r}
img_no_cv <- bn.cv(img_no, bn = bn_img_no, runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
prot_no_cv <- bn.cv(prot_no, bn = bn_prot_no, runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
img_prot_no_cv <- bn.cv(img_prot_no, bn = bn_prot_img_no, runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
img_prot_id_no_cv <- bn.cv(img_prot_no, bn = bn_prot_img_id_no, runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
plot(img_no_cv, prot_no_cv, img_prot_no_cv, img_prot_id_no_cv, xlab = c("Image only", "Proteomics only", "Proteomics -> Image", "Image + Proteomics"))
plot(img_no_cv, img_prot_no_cv, xlab = c("Image only", "Proteomics -> Image"))
```








# Tests

```{r}
img_af_cv <- bn.cv(img_af, bn = "hc", runs = 10, loss = "mse", loss.args = list(target = "affected"))
prot_af_cv <- bn.cv(prot_af, bn = "hc", runs = 10, loss = "mse", loss.args = list(target = "affected"))
img_prot_af_cv <- bn.cv(img_prot_af, bn = "hc", runs = 10, loss = "mse", loss.args = list(target = "affected"))
plot(img_af_cv, prot_af_cv, img_prot_af_cv, xlab = c("Image only", "Proteomics only", "Image + Proteomics"))
```


```{r}
img_no_cv <- bn.cv(img_no, bn = "tabu", runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
prot_no_cv <- bn.cv(prot_no, bn = "tabu", runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
img_prot_no_cv <- bn.cv(img_prot_no, bn = "tabu", runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
plot(img_no_cv, prot_no_cv, img_prot_no_cv, xlab = c("Image only", "Proteomics only", "Image + Proteomics"))
```


```{r}
img_no_cv <- bn.cv(img_no, bn = "hc", runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
prot_no_cv <- bn.cv(prot_no, bn = "hc", runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
img_prot_no_cv <- bn.cv(img_prot_no, bn = "hc", runs = 10, loss = "mse", loss.args = list(target = "NoLN"))
plot(img_no_cv, prot_no_cv, img_prot_no_cv, xlab = c("Image only", "Proteomics only", "Image + Proteomics"))
```


# Using other models to do the cross validation, like AutoML?

```{r}

```



# Hidden Markov Model for modeling the lymphinode dynamics














